---
title: "Generating a Simple, Easily-Shareable Data Dictionary"
author: "Kenneth McLean"
date: "`r Sys.Date()`"
output: md_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE)
```
# Collaborator: Generating a Simple, Easily-Shareable Data Dictionary

The function `data_dict()` can be used to generate an easily sharable and informative data dictionary for an R dataframe. Unlike the `str()` function typically used to display the internal structure of dataframes in R, this produces a dataframe alongside summarising information relevant to the class of variable, and the proportion of missing data (NA) within each variable.

## Requirements
The `data_dict()` function can be applied to any dataframe object. At present, it supports the following classes (other classes will be shown as "Class not supported" in the values column):

* Numeric, integer.
* Logical.
* Date.
* Character, String.
* Factor, orderedfactor.

Example dataframe (`example_data_dict`):
```{r, warning=FALSE, message=FALSE}
library(collaborator)

data <- collaborator::example_data_dict
knitr::kable(head(data, n=5)) # Please note data is not based on real patients

```

## Main Features
### Output
The `data_dict()` function produces a dataframe which identifies the class, summarised values, and proportion of missing data for each variable in the original dataframe.

The output can be easily converted to a csv file and exported for sharing.

```{r, warning=FALSE, message=FALSE}
library(collaborator)
knitr::kable(data_dict(data))

```


Through summarising the variables, data will not necessarily be linkable to individual patients (bar in the circumstance where variable(s) contain a direct patient identifier e.g. Community Health Index (CHI) Number, hospital numbers, etc).

However, should any variable(s) (such as a direct patient identifier) be desirable to exclude from the output, this can be achieved using the "var_exclude" parameter. 

```{r, warning=FALSE, message=FALSE}
library(collaborator)
knitr::kable(data_dict(data, var_exclude = c("id_num","sex")))
```
