---
title: "Generating Authorship Lists"
author: "Kenneth McLean"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE)
  
```
# Collaborator: Generating Authorship Lists

Generating and formatting authorship lists for multi-centre research projects can be a challenging data wrangling task. In the case of collaborative research projects, there can be thousands of collaborators across hundreds of sites with a variety of [roles](https://doi.org/10.1016/j.ijsu.2017.12.019).\n 

The `report_auth()` function aims to simplify this process by providing an easy method to compile a fully formatted authorship list with inbuilt flexibility in how these are presented.

## Requirements
In order for the `report_auth()` function to operate as intended, we must first create a dataframe of all authors/collaborators containing at least 1 column: "name".

Example dataframe (`data_author`):
```{r, warning=FALSE, message=FALSE}
library(collaborator)

data_author <- collaborator::example_report_author
knitr::kable(head(data_author, n=10)) # Please note all names have been randomly generated

```

## Main Features
### (1) Basic Function
At it's most basic, `report_auth()` can produce a formatted list of a column of names.
```{r, warning=FALSE, message=FALSE}

report_auth(data_author) # Please note all names have been randomly generated

```

### (2) Grouping and subdivision of names

These names can be further grouped by another column in the dataframe:
```{r, warning=FALSE, message=FALSE}

report_auth(data_author, group = "hospital")

```

Or can be subdivided by another column in the dataframe:
```{r, warning=FALSE, message=FALSE}

report_auth(data_author, subdivision = "country")

```

Or groups can be further subdivided (for example by region/country, or by role)
```{r, warning=FALSE, message=FALSE}

report_auth(data_author,
            group = "hospital",
            subdivision = "country")

```

### (3) Formatting
Clear and consistent formatting of authorship lists allows the contributions and affiliations of each collaborator/author to be represented. Within `report_auth()`, names are usually separated by a comma (","), with groups separated by a semicolon (";"). Furthermore the name of groups are separated by round brackets ("()"). However, there is a degree of inbuilt flexibility to facilitate customisation.


```{r, warning=FALSE, message=FALSE}

report_auth(data_author, group="hospital", subdivision = "country",
            name_sep = " +", group_brachet = "[}",group_sep = " -")

```
