---
title: "Generating Missing Data Reports"
author: "Kenneth McLean"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Ensuring high levels of completeness within research projects is an important task for ensuring the highest quality dataset for subsequent analyses. However, determining what data is missing within a REDCap project, particularly accounting for appropriately missing data (such as in the case of unfufilled branching logic) can be a challenging and time-consuming task to produce in real-time. 

The `report_miss()` function is designed to easily produce a high quality and informative report of missing data at a data_access_group and individual record level. This report highlights all missing data within a REDCap project (delineating between appropriately missing and true missing data), while removing all other data (so this can be shared in line with duties of data protection).

## Requirements:
The `report_miss()` function is designed to be simple from the point of use - the only requirements are a valid URI (Uniform Resource Identifier) and API (Application Programming Interface) for the REDCap project.

Any variable can be excluded from the output using the "var_exclude" parameter. 

Limitations:
- This function has not yet been tested on REDCap projects with multiple events.

## Output:
### Record level report
Example of a record level report of missing data. This not only quantifies the missing data within the record, but also highlights it's location within the dataset.

- "NA" fields represent appropriately missing data (e.g. secondary to unfufilled branching logic). Therefore, these are excluded from the missing data count entirely.

- "M" fields represent 'true' missing data (which may require follow up), and so all counts of missing data are based these.

```{r, warning=FALSE, message=FALSE}
devtools::install_github("kamclean/collaborator", ref = "development")
library(collaborator)

example_data_miss_record <- data(example_data_miss_record)

```



### Data access group level report
Example of a data access group (DAG) level report of missing data (summarising missing data for all records within the DAG).

```{r, warning=FALSE, message=FALSE}
example_data_miss_dag <- data(example_data_miss_dag)

```
