---
title: "Generating a Simple, Easily-Shareable Data Dictionary"
author: "Kenneth McLean"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(collaborator); library(dplyr)
  
```
# Collaborator: Generating a Simple, Easily-Shareable Data Dictionary

The function `data_dict()` can be used to generate an easily sharable and informative data dictionary for an R dataframe. Unlike the `str()` function typically used to display the internal structure of dataframes in R, this produces a dataframe alongside summarising information relevant to the class of variable, and the proportion of missing data (NA) within each variable.

This can be useful in quickly understanding how data is structured within the dataset, and in assessing data quality (e.g. outlying and incorrect or quantity of missing values). This can be easily exported from R and shared as a spreadsheet.

## Requirements
The `data_dict()` function can be applied to any dataframe object. At present, it supports the following classes (other classes will be shown as "Class not supported" in the values column):

* Numeric, integer.
* Logical.
* Date.
* Character, String.
* Factor, orderedfactor.

Example dataframe (`example_data_dict`):
```{r, warning=FALSE, message=FALSE}
library(collaborator)

data <- collaborator::example_data_dict
knitr::kable(head(data, n=10)) # Please note data is not based on real patients

```

## Main Features
### Output
The `data_dict()` function produces a dataframe which identifies the class, summarised values, and proportion of missing data for each variable in the original dataframe.

The output can be easily converted to a spreadsheet file (e.g. csv file) and exported for sharing.

```{r, warning=FALSE, message=FALSE}

data_dict(data) %>%
  knitr::kable()
  
```


Through summarising the variables, data will not necessarily be linkable to individual patients (bar in the circumstance where variable(s) contain a direct patient identifier e.g. Community Health Index (CHI) Number, hospital numbers, etc).

However, should any variable(s) (such as a direct patient identifier) be desirable to exclude from the output, this can be achieved using the "var_exclude" parameter. 

```{r, warning=FALSE, message=FALSE}
library(collaborator)
knitr::kable(data_dict(data, var_exclude = c("id_num","sex")))
```
