---
title: "Generating REDCap summary data"
author:
 - "Kenneth McLean"
 - "Riinu Ots"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE)

library(collaborator);library(dplyr)
redcap_project_uri = "https://redcap.cir.ed.ac.uk/redcap_v8.4.5/API/"
redcap_project_token = "582D6629348CB3AED679BCADE20969DA"

```
# Collaborator: REDCap summary data
Evaluating data uploaded to a REDCap project in the context of multi-centre research projects is an important task for ensuring the highest quality dataset for subsequent analyses, and for sharing progress (whether internally or externally via social media). However, summarising this data, particularly at a DAG-level, can be a challenging and time-consuming task to produce on a regular basis.

The `redcap_sum()` function is designed to easily produce high quality and informative summary data on a REDCap project (overall and at a data_access_group level). This can be used for the purposes of sharing progress (including identifying top performing DAGs), and identifying individual DAGs which have not yet uploaded data or have not completed data upload.

## Requirements
The `redcap_sum()` function is designed to be simple from the point of use - the only requirements are a valid URI (Uniform Resource Identifier) and API (Application Programming Interface) for the REDCap project.

However, this is intended to have a high degree of customisability to fit the needs of a variety of projects. For example, being able to easily:

- Exclude individual records (`record_exclude`) or whole data access groups (`dag_exclude`) from the record count (e.g. records or DAGs that were found to be ineligible).

- Exclude users (`user_exclude`) from the total REDCap user count (e.g. administrator user accounts).

- Define variables that should contribute towards evaluation of data completeness. This can be achieved by either excluding  (`var_exclude`), or specifying certain variables (`var_complete`).

- Generation of summary data by DAG unless `centre_sum` is specified as FALSE (default `centre_sum=T`) 

Limitations:

- This function has not yet been tested on REDCap projects with multiple events.

## Main Features
### (1) Basic Function
At it's most basic, `redcap_sum()` can produce an overall summary of current data on the REDCap project:
- `n_record_all` is the number of all records currently on the REDCap project (minus any records removed using `record_exclude` or in DAGs removed using `dag_exclude`).
 
- `n_record_com` is the number of all complete records currently on the REDCap project (minus any records removed using `record_exclude` or in DAGs removed using `dag_exclude`), with no missing data across the record (unless certain data fields are either excluded (`var_exclude`) or specified (`var_complete`)).
 
- `prop_com` /`pct_com` is the respective proportion and percentage of complete records in the project.
 
- `n_dag` is the number of data access groups (DAGs) for all records currently on the REDCap project (minus any records removed using `record_exclude` or in DAGs removed using `dag_exclude`).
 
- `n_users` is the number of users on the REDCap project (minus any users in DAGs removed using `dag_exclude`). Note all users not assigned to a DAG will automatically be excluded.
 
- `last_update` is the date which the summary data was generated on. 
 
```{r, warning=FALSE, message=FALSE}
redcap_sum(redcap_project_uri, redcap_project_token, centre_sum = F)  %>%
  knitr::kable()

```


### (2) Centre summary data
However, more granular summary data on each DAG can also be obtained using the same function. This centre summary data will automatically be included within the output from `redcap_sum()` unless `centre_sum` is specified as FALSE. 

#### 1. `$dag_all` Output
This will produce a dataframe of the same summary data as outlined above **grouped by each DAG instead** (minus any DAGs removed using `dag_exclude`).

```{r, warning=FALSE, message=FALSE}
redcap_sum(redcap_project_uri, redcap_project_token, centre_sum = T)$dag_all %>%
  knitr::kable()

```

#### 2. `$dag_nodata` Output
This will produce a dataframe of all DAG with users assigned on the REDCap project, but no data uploaded to REDCap. This may be useful for the purposes of targeting encouragement to upload data, or establishing authorship on any research output. 

```{r, warning=FALSE, message=FALSE}
redcap_sum(redcap_project_uri, redcap_project_token, centre_sum = T)$dag_nodata %>%
  knitr::kable()

```


#### 3. `$dag_incom` Output
This will produce a dataframe of all DAG with incomplete records (the definition of completeness customisable as discussed above). This may be useful for the purposes of follow up regarding (essential) missing data at each of these DAGs. 

```{r, warning=FALSE, message=FALSE}
redcap_sum(redcap_project_uri, redcap_project_token, centre_sum = T)$dag_incom %>%
  knitr::kable()

```



#### 4. `$dag_top_n` Output
This will produce a dataframe of the DAGs with the most records uploaded overall (the number of DAGs listed is defined by `n_top_dag` with top 10 DAG being default). This may be useful for the purposes of publicity surrounding the project. 


```{r, warning=FALSE, message=FALSE}
redcap_sum(redcap_project_uri, redcap_project_token, centre_sum = T, n_top_dag = 5)$dag_top_n %>%
  knitr::kable()

```
