---
title: "Generating a Simple, Easily-shareable Data Dictionary"
author: "Kenneth McLean"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The function `data_dict()` can be used to generate an easily sharable and informative data dictionary for an R dataframe. Unlike the `str()` function typically used to display the internal structure of dataframes in R, this produces a dataframe alongside summarising information relevent to the class of variable, and the proportion of missing data (NA) within each variable.

## Requirements
The `data_dict()` function can be applied to any dataframe object. At present, it supports the following classes (other classes will be shown as "Class not supported" in the values column):
- Numeric, integer.
- Logical.
- Date.
- Character, String.
- Factor, orderedfactor.
example_data_dict <- 
Example dataframe (`example_data_dict`):
```{r, warning=FALSE, message=FALSE}
devtools::install_github("kamclean/collaborator", ref = "development")
library(collaborator)

example_data_dict <- collaborator::example_data_dict
head(example_data_dict, n=5) # Please note data is not based on real patients

```

## Main Features
### Output
The `data_dict()` function produces a dataframe which identifies the class, summarised values, and proportion of missing data for each variable in the original dataframe.

The output can be easily converted to a csv file and exported for sharing.

```{r, warning=FALSE, message=FALSE}
library(collaborator)
data_dict(example_data_dict)

```


Any variable can be excluded from the output using the "var_exclude" parameter. Through summarising the variables, data should not be linkable to individual patients (bar in the circumstance where variable(s) contain a direct patient identifer e.g. Community Health Index (CHI) Number, hospital numbers, etc).

```{r, warning=FALSE, message=FALSE}
library(collaborator)
data_dict(example_data_dict, var_exclude = "id_num")
```
